---
title: "Kenya COVID-19 Vaccination Coverage"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: embed
    vertical_layout: fill
    self_contained: FALSE
---
```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(sf)
library(RecordLinkage)
library(lubridate)
# imported data -----------------------------------------------------------

data<-read_csv("vaccination.csv")
counties<-st_read("County.shp")
vaccination_data<-read_csv("covid19_vaccine_data_20210422.csv")
facilities<- read_csv("facilities.csv")%>%
  select(Officialname, County)%>%
  distinct()%>%
  mutate(Officialname=str_to_title(Officialname))

# data wrangling ----------------------------------------------------------

vaccination_data1<- vaccination_data%>%
  mutate(vaccination_date=date(vaccination_date))%>%
  mutate(place_of_vaccination=str_to_title(place_of_vaccination))%>%
  left_join(facilities, by=c("place_of_vaccination"="Officialname"))%>%
  mutate(County=ifelse(place_of_vaccination%in%c("Aga Khan Hospital Mombasa", "Mombasa Regional Vaccines Store", 
                                                 "Tudor Sub-County Hospital", "Likoni Sub-County Hospital", 
                                                 "Port Reitz  Hospital", "Premiere Hospital"), "Mombasa", 
                ifelse(place_of_vaccination%in%c("Alupe Sub County Hospital", "Holly Family Nangina Mission Hospital", 
                                                 "Nambale Sub County", "Port Victoria Sub County Hospital"), "Busia",
                ifelse(place_of_vaccination%in%c("Baragoi Sub- County Hospital", "Samburu Sub County Hospital"), "Samburu",
                ifelse(place_of_vaccination%in%c("Yala Sch", "Ukwala Sch"), "Siaya",
                ifelse(place_of_vaccination%in%c("Kutulo Sub County Hospital", "Mandera North Subcounty Hospital", "Lafey Sub County Hospital"), "Mandera",
                ifelse(place_of_vaccination%in%c("Athi River Level 4", "Machakos Sub County", "Kangundo Level 4","Kathiani Level 4", 
                                                 "Masinga Level 4", "Masinga Sub County", "Matuu Level 4", "Muuma Andu Level 4", 
                                                 "Mwala Level 4", "Mutituni Level 4", "Nguluni Level 4"), "Machakos", 
                ifelse(place_of_vaccination%in%c("Aic Litein Hospital", "Fort Ternan Sch", "Kapkatet Sub County Hospital", "Kipkelion Sub County Hospital",
                                                 "Kipchimchim Hospital", "Londiani Hospital", "St. Gabriel Hospital", "Sosiot Sch",
                                                 "Roret Sub County Hospital", "Sigowet Sub County Hospiotal", "Sigowet/Soin Sub County"), "Kericho", 
                ifelse(place_of_vaccination%in%c("Awendo Subcounty Hospital", "Isebania Subcounty Hospital", 
                                                 "Karungu Subcounty Hospital", "Kegonga Subcounty Hospital", "Kehancha Subcounty Hospital", 
                                                 "Kuria East Sub County", "Kuria West Sub County","Macalder Subcounty Hospital", 
                                                 "Uriri Subcounty Hospital", "Ulanda Mission Hospital", "Lifecare Hospital",
                                                 "Muhuru Sub County Hospital", "Nyatike Sub County", "Pastor Machaje Hospital", 
                                                 "Othoro Subcounty Hospital", "St. Josephs Mission", "St. Camillus Sori", 
                                                 "Oyani Subcounty Hospital", "Rapcom Hospital", "Rongo Sub County", 
                                                 "Rongo Subcounty Hospital", "Royal Hospital"), "Migori",
                ifelse(place_of_vaccination%in%c("Bliss Healthcare, College House", "Bliss Healthcare, Embakasi",
                                                 "Bliss Healthcare, Westlands", "Coptic Mission Hospital", 
                                                 "Dagoretti North Sub County", "Nairobi Regional Vaccines Store",
                                                 "Afya House Dispensary", "Houghton Medical Centre", "National Distribution Center", 
                                                 "Quiver Medical Center"), "Nairobi",
                ifelse(place_of_vaccination%in%c("Bomet Central Sub County", "Bomet East Sub County", "Chepalungu Sub County", 
                                                 "Konoin Sub County", "Sotik Sub County"), "Bomet",
                ifelse(place_of_vaccination%in%c("Bokoli Sub-County Hospital", "Bumula Sub-County Hospital", 
                                                 "Bungoma West Hospital", "Cheptais Sub County", 
                                                 "Cheptais Sub-County Hospital", "Chwele Sub-County Hospital", "Dreamland Medical Cenre",
                                                 "Khalaba Medical Centre", "Kimilili Sub County","Kimilili Sub-County Hospital",
                                                 "Korry Family Hospital", "Webuye County Hospital", "Lugulu Mission Hospital", 
                                                 "Mt. Elgon Sub-County Hospital", "Naitiri Sub-County Hospital", 
                                                 "Sinoko Sub-County Hospital","Sirisia Sub-County Hospital"), "Bungoma",
                ifelse(place_of_vaccination%in%c("Bura Sub County Hospital"), "Tana River",
                ifelse(place_of_vaccination%in%c("Mwatate Sub County Hospital"), "Taita Taveta",
                ifelse(place_of_vaccination%in%c("Chemolingot District Hospital", "Butere County Hospital", "Iguhu County Hospital", 
                                                 "Khwiseo Health Centre", "Likuyani Sub County Hospital", "Lumakanda County Hospital",
                                                 "Malava County Hospital", "Mumias West Sub County Hospital", 
                                                 "Navakholo Sub County Hospital", "Shianda Health Centre"), "Kakamega",
                ifelse(place_of_vaccination%in%c("Banissa Sub County Hospital", "Eldama Ravine District Hospital", "Mogotio  Sub-County Hospital Baringo"), "Baringo", 
                ifelse(place_of_vaccination%in%c("Chepareria Sub County  Hospital", "Kabichbich Health  Centre",
                                                 "Kacheliba Sub County Hospital", "Sigor Sub Couunty  Hospital"), "West Pokot",
                ifelse(place_of_vaccination%in%c("Chepterwai Sub County Hospital", "Kaptumo Sub County Hospital", "Meteitei Sub County Hospital"), "Nandi",
                ifelse(place_of_vaccination%in%c("Engineer County Hospital"), "Nyandarua",
                ifelse(place_of_vaccination%in%c("Huruma Sub County Hospital", "Fountain Health Care", "Kapteldon Sub-County Hospital", 
                                                 "Kesses Sub-County Hospital", "Mediheal Hospital-Eldoret", "Moiben Sub County Hospital"), "Uasin Gishu",
                ifelse(place_of_vaccination%in%c("Faza Sub County Hospital", "Mpeketoni Sub County Hospital"), "Lamu",
                ifelse(place_of_vaccination%in%c("Loitokitok Sub-County Hospital", "Ongata Rongai Sub-County Hospital"), "Kajiado",
                ifelse(place_of_vaccination%in%c("Kakuma Sub County Hospital", "Turkana West Sub County", "Lokichar Sub County Hospital",
                                                 "Lokori Sub County Hospital", "Lorugum Sub County Hospital"), "Turkana",
                ifelse(place_of_vaccination%in%c("Gesima Sub-County Hospital", "Masaba North Sub County", "Nyamira North Sub County"), "Nyamira",
                ifelse(place_of_vaccination%in%c("Ijara District Hospital - Masalani", "Modogashe Sub-County Hospital", "Msf Dagahaley Main Hospital"), "Garissa",
                ifelse(place_of_vaccination%in%c("Kabondo Sub-County  Hospital", "Marindi Sub County Referral Hospital", "Mbita Sub County", 
                                                 "Ndhiwa Sub-District Hospital", "Suba Sub-Couty Hospital", "Rachuonyo South Sub County"), "Homa Bay",
                ifelse(place_of_vaccination%in%c("Ishiara Level 4 Hospital","Mbeere Level 4 Hospital"), "Embu",
                ifelse(place_of_vaccination%in%c("Kinondo Hospital", "Lunga Lunga Sub County Hospital", "Matuga Sub County"), "Kwale",
                ifelse(place_of_vaccination%in%c("Wangige Sub County Hospital", "Lari Hospital", "Medicross Limited, Thika Ananas"), "Kiambu",
                ifelse(grepl("Igembe|Tigania", place_of_vaccination), "Meru",
                ifelse(place_of_vaccination%in%c("Family Heritage  Hospital", "Kangema Sub County Hospital", "Kigumo Level 4 Hospital", 
                                                 "Kigumo Sub County Hospital", "Kirwara Sub County Hospital", "Maragua Sub County Hospital", 
                                                 "Muranga", "Muriranjas Sub County Hospital"), "Murang'a",
                ifelse(place_of_vaccination%in%c("Eldas Sub County", "Griftu Sub County Hospital", "Habaswein Sub County Hospital", 
                                                 "Wajir North Sub County", "Tarbaj Sub County Hospital"), "Wajir",
                ifelse(place_of_vaccination%in%c("Emuhaya Sub-County Hospital", "Hamisi Sub-County Hospital", "Kaimosi Mission", 
                                                 "Sabatia Sub-County Hospital"), "Vihiga",
                ifelse(place_of_vaccination%in%c("Cherangany Sub-County Hospital", "Endebess Sub-County Hospital",
                                                 "Kapsara Sub-County Hospital", "Kwanza Sub-County Hospital"), "Trans Nzoia",
                ifelse(place_of_vaccination%in%c("Chuka Sub County", "Magutuni Sub-County Hospital", "Pcea Chogoria Hospital"), "Tharaka Nithi",
                ifelse(place_of_vaccination%in%c("Consolata Mission Hospital-Mathari", "Nyeri Regional Vaccines Store",
                                                 "Narumoru Health Centre", "Kaiyaba Health Centre", "Karatina Sub-County Hospital", 
                                                 "Tumutumu Hospital", " Mathira East Sub County", "Mt.kenya Hospital", "Mukurweini Hospital",
                                                 "Mathira East Sub County", "Mathira West Sub County", "Othaya Sub-County Hospital"), "Nyeri",
                ifelse(grepl("Kisumu|Muhoroni|Nyando|Nyakach|Seme Sub", place_of_vaccination), "Kisumu",
                ifelse(grepl("Kitui|Yatta|Mwingi|Mutomo", place_of_vaccination), "Kitui",
                ifelse(grepl("Nanyuki|Nyahururu", place_of_vaccination), "Laikipia",
                ifelse(place_of_vaccination%in%c("Marakwet West Sub County", "Marakwet East Sub County", "Keiyo North Sub County", 
                                                 "Keiyo North Sub County", "Keiyo South Sub County"), "Elgeyo Marakwet",
                ifelse(place_of_vaccination%in%c("Etago Sub County Hospital", "Gesusu Sub County Refferal Hospital", "Iranda Sub County Referral Hospital",
                                                 "Iyabe Sub County Referral Hospital", "Kenyenya Sub Count Referral Hospital", 
                                                 "Keumbu Sub County Referral Hospital", "Kitutu Chache South Sub County",
                                                 "Marani Sub County Referral Hospital","Nyamache Sub County Refferal Hospital",
                                                 "St Cathrine Ichuni Mission Hospital"), "Kisii",
                ifelse(place_of_vaccination%in%c("Laisamis Sub County Referral Hospital(Marsabit South)","Ismc Hospital", 
                                                 "Moyale Sub County Referral  Hospital"), "Marsabit",
                ifelse(grepl("Nakuru|Gilgil Sub|Keringet|Kuresoi|Lanet|Langa Langa|Molo|Njoro|St. Elizabeth|Rongai Sub County", place_of_vaccination), "Nakuru",
                ifelse(grepl("Kilifi|Ganze Sub County|Magarini|Malindi|Rabai Sub", place_of_vaccination), "Kilifi",
                ifelse(place_of_vaccination%in%c("Sultan Hamud Subcounty Hospital (Level Iv)", "Tawa Subcounty Hospital (Level Iv)",
                                                 "Kibwezi Subcounty Hospital (Iv)", "Makindu Subcounty Hospital (Level Iv)",
                                                 "Kambu Subcounty Hospital (Iv)", "Kikoko Mission Hospital (Iv)",
                                                 "Kilungu Subcounty Hospital (Iv)", "Kisau Subcounty Hospital (Level Iv)", 
                                                 "Matiliku Subcounty Hospital (Level Iv)", "Mbooni Subcounty Hospital (Level Iv)", 
                                                 "Mtito Andei Subcounty Hospital (Iv)", "Mukuyuni Subcounty Hospital (Level Iv)", 
                                                 "Mutyambua Subcounty Hospital (Level Iv)", "Mulatya Memorial Hospital", 
                                                 "Ralpha Hospital", "Royal Star Hospital"), "Makueni",County 
                       )))))))))))))))))))))))))))))))))))))))))))))%>%
  mutate(County=ifelse(is.na(County), place_of_vaccination, County))%>%
  mutate(County=recode(County, "Muranga"="Murang'a"))

```

Summary {data-navmenu="Summary"}
============================================================

```{r, echo=FALSE,message=F, warning=F, fig.width=16, fig.height=10}

vaccination_data1$County<-recode(vaccination_data1$County, "Murang'a"="Muranga")
vaccination_data1a<-vaccination_data1%>%
  group_by(County)%>%
  count()%>%
  ungroup()
counties1<- full_join(counties, vaccination_data1a, by=c("Name"="County"))

p1<-ggplot(counties1, aes(fill=n))+geom_sf()+theme_bw()+scale_fill_gradient(low="#ffeda0", high="#f03b20")+labs(fill="Number of people \nvaccinated")

p2<-ggplot(counties1[counties1$Name!="Nairobi",], aes(fill=n))+geom_sf()+theme_bw()+scale_fill_gradient(low="#ffeda0", high="#f03b20", na.value = "white")+labs(fill="Number of people \nvaccinated", caption="Nairobi has been omitted for clarity")

gridExtra::grid.arrange(p1, p2, nrow=1)
```

- The vaccination dates were ranging from 1994 to 2029. 

- In plotting this graphic, we used data for 2021 though some entries were from 1st January 2021 yet the first batch of the vaccine arrived in Kenya on 2nd March 2021.

Trends {data-navmenu="Trends"}
============================================================

```{r, echo=F,message=F, warning=F, fig.width=16, fig.height=10}
vaccination_data1b<- vaccination_data1%>%
  filter(vaccination_date>as.Date("2021-01-01") & vaccination_date<as.Date("2021-04-29"))%>%
  group_by(vaccination_date)%>%
  count()%>%
  ungroup()


ggplot(vaccination_data1b, aes(x=vaccination_date, y=n))+geom_col(fill="darkgreen")+theme_bw()+scale_x_date(breaks="1 month", date_labels = "%b-%y")+labs(x="Period", y="Number of people vaccinated")
```

By occupation {data-navmenu="By Occupation"}
============================================================

```{r, echo=F,message=F, warning=F, fig.width=16, fig.height=10}
vaccination_data1c<- vaccination_data1%>%
  group_by(age)%>%
  count()%>%
  ungroup()

ggplot(vaccination_data1c[vaccination_data1c$age>0,], aes(x=age, y=n))+geom_col(fill="darkgreen")+theme_bw()+labs(x="Age (Years)", y="Number of people vaccinated")+scale_x_continuous(breaks = seq(18,99,by=10))+geom_vline(xintercept = 58, linetype=2, color="red")
```
```{r, echo=F,message=F, warning=F}
##occupation
library(vcd)
library(ggmosaic)
vaccination_data1d<- vaccination_data1%>%
  group_by(occupation, County)%>%
  count()%>%
  ungroup()


# ggplot(data = vaccination_data1) +
#   geom_mosaic(aes(x = product(occupation,County), fill = County), show.legend = F)+
#   theme_bw()

#mosaic(Freq~occupation+County, data=vaccination_data1, shade=T, legend=T)
ggplot(vaccination_data1d, aes(x=reorder(County,n),y=n, fill=occupation))+geom_col()+coord_flip()+theme_bw()+labs(y="Numberof people vaccinated", x="County", fill="Occupation")
#mosaicplot(t(subs_count), col = vaccination_data1, off = 5, las = 1, main = "")
```

By type {data-navmenu="By Type"}
============================================================

```{r, echo=F,message=F, warning=F, fig.width=16, fig.height=10}
## type of vaccination

vaccination_data1e<- vaccination_data1%>%
  group_by(vaccine, County)%>%
  count()%>%
  ungroup()


ggplot(vaccination_data1e, aes(x=reorder(County,n),y=n, fill=vaccine))+geom_col()+coord_flip()+theme_bw()+labs(y="Numberof people vaccinated", x="County", fill="Type of vaccine")
```

- Almost all the people vaccination got the Covidshield vaccine (100%, n=780025) while less than 1% got the Sputnik-V vaccine (n=1296)

- Out of the people vaccinated, 13% (n=101,164) had an underlying condition.

Underlying conditions {data-navmenu="Underlying conditions"}
============================================================

```{r, echo=F, message=F, warning=F, fig.width=16, fig.height=10}
## underlying conditions
library(splitstackshape) 
vaccination_data1f<- cSplit(vaccination_data1, 'conditions', ',')

vaccination_data1f<-reshape2::melt(vaccination_data1f, id=1:8)

vaccination_data1g<- vaccination_data1f%>%
  mutate(value=str_remove(value, pattern="\\["))%>%
  mutate(value=str_remove(value, pattern="\\]"))%>%
  mutate(value=str_remove(value, pattern="\\'"))%>%
  mutate(value=str_remove(value, pattern="\\'"))%>%
  mutate(value=trimws(value))%>%
  mutate(value=ifelse(value%in%"", NA, value))%>%
  select(-variable)%>%
  distinct()%>%
  filter(!is.na(value))%>%
  group_by(value, County)%>%
  count()

vaccination_data1h<- vaccination_data1f%>%
  mutate(value=str_remove(value, pattern="\\["))%>%
  mutate(value=str_remove(value, pattern="\\]"))%>%
  mutate(value=str_remove(value, pattern="\\'"))%>%
  mutate(value=str_remove(value, pattern="\\'"))%>%
  mutate(value=trimws(value))%>%
  mutate(value=ifelse(value%in%"", NA, value))%>%
  select(-variable)%>%
  distinct()%>%
  filter(!is.na(value))%>%
  group_by(value)%>%
  count()

ggplot(vaccination_data1h, aes(x=reorder(value, n), y=n))+geom_col(fill="#feb24c")+theme_bw()+coord_flip()+labs(x="Pre-existing condition", y="Number vaccinated")


```

